Type cones of $\b{g}$-vector fans

### 2019-02-08 thirdPaper/draft/typeConeAssociahedra.tex
Removed empty Section 3, added intro to new section 3

% geometry
\newcommandx{\Asso}[2][1=n,2={}]{\mathsf{Asso}^{#2}(#1)} % associahedron
\newcommandx{\Zono}[2][1=n,2={}]{\mathsf{Zono}^{#2}(#1)} % zonotope
\newcommand{\Fan}{\mathcal{F}} % fan
\newcommand{\multiplicityVector}{\b{m}} % multiplicity vector
\newcommandx{\ray}[1][1=r]{\b{#1}} % ray
\newcommandx{\rays}[1][1=R]{\b{#1}} % rays
\newcommand{\gvector}[1]{\b{g}(#1)} % g-vector of #1
\newcommand{\gvectorFull}[2]{\b{g}_{#1}(#2)} % g-vector of #2 wrt #1
\newcommand{\gvectors}[1]{\b{g}(#1)} % g-vectors of #1
\newcommand{\gvectorsFull}[2]{\b{g}_{#1}(#2)} % g-vectors of #2 wrt #1
\newcommandx{\gvectorFan}[1][1=\quiver]{\mathcal{F}(#1)} % g-vector fan
\newcommand{\cvector}[2]{\mathbf{c}(#2 \in #1)} % c-vector of the cluster variable #2 in the cluster #1
\newcommand{\cvectorFull}[3]{\mathbf{c}_{#1}(#3 \in #2)} % c-vector of the cluster variable #3 in the cluster #2 with respect to the initial cluster #1
\newcommand{\cvectors}[1]{\mathbf{c}(#1)} % c-vectors of the cluster #1
\newcommand{\cvectorsFull}[2]{\mathbf{c}_{#1}(#2)} % c-vectors of the cluster #2 with respect to the initial cluster #1

@@

% Cluster algebras
\newcommand{\Trop}[1]{\mathbb{P}_{#1}} % tropical semifield
\newcommand{\positiveExponents}[1]{\left\{#1\right\}_+} % positive exponents in tropical semifield
\newcommand{\negativeExponents}[1]{\left\{#1\right\}_-} % negative exponents in tropical semifield
\newcommand{\seed}{\Sigma} % a cluster
\newcommand{\cluster}{\mathrm{X}} % a cluster
\newcommandx{\clusters}{\mathcal{X}} % all clusters
\newcommand{\coefficients}{\mathrm{P}} % a coefficients tuple
\newcommandx{\variables}[1][1=\clusterAlgebra]{\mathcal{V}(#1)} % all variables
\newcommandx{\clusterAlgebra}[2][1=\B_\circ, 2=\coefficients_\circ]{\mathcal{A}\!\left(#1,#2\right)} % cluster algebra
\newcommandx{\principalClusterAlgebra}[1][1=\B_\circ]{\mathcal{A}_{\mathrm{pr}}\left(#1\right)} % principal coefficients cluster algebra
\newcommandx{\coefficientFreeClusterAlgebra}[1][1=\B_\circ]{\mathcal{A}_{\mathrm{fr}}(#1)} % coefficient free cluster algebra
\newcommandx{\clusterComplex}[1][1=\B_\circ]{\Delta(#1)} % cluster complex
\newcommand{\B}{\mathrm{B}} % b-matrix
\newcommand{\A}[1]{\mathrm{A}({#1})} % b-matrix
\newcommand{\D}{\mathrm{D}} % symmetrizer
\newcommand{\simpleRoot}{\alpha} % simple root
\newcommand{\fundamentalWeight}{\omega} % fundamental weights

@@

\subsubsection{Cluster algebras and $\b{g}$-vector fans}

We present some definitions and properties of finite type cluster algebras and their $\b{g}$-vector fans, following the presentation of~\cite{HohlwegPilaudStella}.

@@

\begin{definition}[\cite{FominZelevinsky-ClusterAlgebrasIV}]
The \defn{$\b{g}$-vector}~$\gvector{x} = \gvectorFull{\B_\circ}(x)$ of a cluster variable~$x \in \principalClusterAlgebra$ is its degree.
We denote by $\gvectors{\seed} = \gvectorFull{\B_\circ}(\seed) \eqdef \set{\gvectorFull{\B_\circ}(x)}{x \in \seed}$ the set of $\b{g}$-vectors of the cluster variable in the seed~$\seed$ of~$\principalClusterAlgebra$.
\end{definition}

@@

\begin{definition}
Given a seed~$\seed$ in~$\principalClusterAlgebra$, the \defn{$\b{c}$-vector} of a cluster variable $x \in \seed$ is the vector
\(
\cvector{\seed}{x} = \cvectorFull{\B_\circ}{\seed}{x} \eqdef \sum_{y \in \cluster_\circ} c_{yx} \, \simpleRoot_y
\)
of exponents of~$p_x = \prod_{y \in \cluster_\circ} (p_y)^{c_{yx}}$.
We denote by~$\cvectors{\seed} = \cvectorsFull{\B_\circ}{\seed} \eqdef \set{\cvectorFull{\B_\circ}{\seed}{x}}{x \in \seed}$ the set of $\b{c}$-vectors of a seed~$\seed$.
\end{definition}

@@

\begin{theorem}[{\cite[Theorem~1.2]{NakanishiZelevinsky}}]
\label{thm:duality}
For any seed~$\seed$ of~$\principalClusterAlgebra$, let $\seed^\vee$ be its dual in $\principalClusterAlgebra[\B_\circ^\vee]$.
Then the set of $\b{g}$-vectors~$\gvectorFull{\B_\circ}{\seed}$ and the set of $\b{c}$-vectors~$\cvectorsFull{\B_\circ^\vee}{\seed^\vee}$ form dual bases, that is
\(
{\bigdotprod{\gvectorFull{\B_\circ}{x}}{\cvectorFull{\B_\circ^\vee}{\seed^\vee}{y^\vee}} = \delta_{x=y}}
\)
for any two cluster variables~$x,y \in \seed$.
\end{theorem}

\para{The $\b{g}$-vector fan and generalized associahedron}
%
The following statement is well known.
\vincent{Explain}

@@

Moreover, this fan is known to be polytopal.
More precisely, consider a vector~$\b{h} \in \R^{\variables}$ such that
\[
\b{h}_x + \b{h}_{x'} > \max \Big( \sum_{{y \in \cluster \cap \cluster', \; b_{xy} < 0}} -b_{xy} \, \b{h}_y \;,  \sum_{{y \in \cluster \cap \cluster', \; b_{xy} > 0}} b_{xy} \, \b{h}_y \Big).
\]
for any adjacent seeds~$(\B, \coefficients, \cluster)$ and~$(\B', \coefficients', \cluster')$ with~$\cluster \ssm \{x\} = \cluster' \ssm \{x'\}$.
Such a vector~$\b{h}$ exists, see the discussion in~\cite[Prop.~28]{HohlwegPilaudStella}.

\begin{theorem}[{\cite[Thm.~26]{HohlwegPilaudStella}}]
\label{thm:gvectorFan}
For any finite type exchange matrix~$\B_\circ$, the $\b{g}$-vector fan~$\gvectorFan[\B_\circ]$ is the normal fan of the \defn{$\B_\circ$-associahedron}~$\Asso{\B_\circ, \b{h}}$ defined equivalently as
\begin{enumerate}[(i)]
\item the convex hull of the points~$\sum_{x \in \seed} \b{h}_x \, \cvectorFull{\B_\circ^\vee}{\seed^\vee}{x^\vee}$ for all seeds~$\seed$ of~$\principalClusterAlgebra$, or
\item the intersection of the halfspaces~$\set{\b{v} \in V^\vee}{\dotprod{\gvectorFull{\B_\circ}{x}}{\b{v}} \le \b{h}_x}$ for all cluster variables~$x$ of~$\principalClusterAlgebra$.
\end{enumerate}
\end{theorem}

@@

\begin{remark}
\label{rem:exchangeablePairsNKC}
In view of \cref{prop:exchangeablePairsNKC}, it is tempting to look for a characterization of the exchangeable pairs~$\omega, \omega'$ using the kisses between~$\omega$ and~$\omega'$.
However, as illustrated in \cref{fig:nonkissingFans}\,(right), note that
\begin{itemize}
\item two exchangeable walks may kiss along more than one string, but only one is the distinguished string,
\item two non-exchangeable walks can kiss along more than one distinguishable string,
\item two walks that kiss along a single distinguishable string are not always exchangeable,
\item not all strings are distinguishable.
\end{itemize}
\vincent{More explicit. TODO}
\vincent{Do we still have these problems when the gentle algebra is brick and/or without oriented $2$-cycle?}
\end{remark}

@@

\section{Relations for~$\b{g}$-vectors in~$2$-Calabi--Yau triangulated categories}

@@

@@

%%
\subsubsection{$2$-term complexes of projective modules}

@@

%%
\subsubsection{Associative algebras}\label{sect::Kbproj}

>>>>

% geometry
\newcommandx{\Asso}[2][1=n,2={}]{\mathsf{Asso}^{#2}(#1)} % associahedron
\newcommandx{\Zono}[2][1=n,2={}]{\mathsf{Zono}^{#2}(#1)} % zonotope
\newcommand{\Fan}{\mathcal{F}} % fan
\newcommand{\multiplicityVector}{\b{m}} % multiplicity vector
\newcommandx{\ray}[1][1=r]{\b{#1}} % ray
\newcommandx{\rays}[1][1=R]{\b{#1}} % rays
\newcommand{\gvector}[1]{\b{g}(#1)} % g-vector of the path #1
\newcommand{\gvectors}[1]{\b{g}(#1)} % g-vectors of the paths #1
\newcommandx{\gvectorFan}[1][1=\quiver]{\mathcal{F}(#1)} % g-vector fan
\newcommand{\cvector}[3]{\mathbf{c}(#1, #3 \in #2)} % c-vector of the cluster variable #3 in the cluster #2 with respect to the initial cluster #1
\newcommand{\cvectors}[2]{\mathbf{c}(#1, #2)} % c-vectors of the cluster #2 with respect to the initial cluster #1

@@

% Cluster algebras
\newcommand{\Trop}[1]{\mathbb{P}_{#1}} % tropical semifield
\newcommand{\positiveExponents}[1]{\left\{#1\right\}_+} % positive exponents in tropical semifield
\newcommand{\negativeExponents}[1]{\left\{#1\right\}_-} % negative exponents in tropical semifield
\newcommand{\seed}{\Sigma} % a cluster
\newcommand{\cluster}{\mathrm{X}} % a cluster
\newcommandx{\clusters}{\mathcal{X}} % all clusters
\newcommand{\coefficients}{\mathrm{P}} % a coefficients tuple
\newcommandx{\clusterAlgebra}[2][1=\B_\circ, 2=\coefficients_\circ]{\mathcal{A}\!\left(#1,#2\right)} % cluster algebra
\newcommandx{\principalClusterAlgebra}[1][1=\B_\circ]{\mathcal{A}_{\mathrm{pr}}\left(#1\right)} % principal coefficients cluster algebra
\newcommandx{\coefficientFreeClusterAlgebra}[1][1=\B_\circ]{\mathcal{A}_{\mathrm{fr}}(#1)} % coefficient free cluster algebra
\newcommandx{\clusterComplex}[1][1=\B_\circ]{\Delta(#1)} % cluster complex
\newcommand{\B}{\mathrm{B}} % b-matrix
\newcommand{\A}[1]{\mathrm{A}({#1})} % b-matrix
\newcommand{\D}{\mathrm{D}} % symmetrizer
\newcommand{\simpleRoot}{\alpha} % simple root
\newcommand{\fundamentalWeight}{\omega} % fundamental weights

@@

@@

\begin{definition}[\cite{FominZelevinsky-ClusterAlgebrasIV}]
The \defn{$\b{g}$-vector}~$\gvector{\B_\circ}{x}$ of a cluster variable~$x \in \principalClusterAlgebra$ is its degree.
We denote by $\gvectors{\B_\circ}{\seed} \eqdef \set{\gvector{\B_\circ}{x}}{x \in \seed}$ the set of $\b{g}$-vectors of the cluster variable in the seed~$\seed$ of~$\principalClusterAlgebra$.
\end{definition}

@@

\begin{definition}
Given a seed~$\seed$ in~$\principalClusterAlgebra$, the \defn{$\b{c}$-vector} of a cluster variable $x \in \seed$ is the vector
\(
\cvector{\B_\circ}{\seed}{x} \eqdef \sum_{y \in \cluster_\circ} c_{yx} \, \simpleRoot_y
\)
of exponents of~$p_x = \prod_{y \in \cluster_\circ} (p_y)^{c_{yx}}$.
We denote by~$\cvectors{\B_\circ}{\seed} \eqdef \set{\cvector{\B_\circ}{\seed}{x}}{x \in \seed}$ the set of $\b{c}$-vectors of a seed~$\seed$.
\end{definition}

@@

\begin{theorem}[{\cite[Theorem~1.2]{NakanishiZelevinsky}}]
\label{thm:duality}
For any seed~$\seed$ of~$\principalClusterAlgebra$, let $\seed^\vee$ be its dual in $\principalClusterAlgebra[\B_\circ^\vee]$.
Then the set of $\b{g}$-vectors~$\gvector{\B_\circ}{\seed}$ and the set of $\b{c}$-vectors~$\cvectors{\B_\circ^\vee}{\seed^\vee}$ form dual bases, that is
\(
{\bigdotprod{\gvector{\B_\circ}{x}}{\cvector{\B_\circ^\vee}{\seed^\vee}{y^\vee}} = \delta_{x=y}}
\)
for any two cluster variables~$x,y \in \seed$.
\end{theorem}

\para{The $\b{g}$-vector fan}
%
The following statement is well known.
\vincent{Explain}

@@

Moreover, this fan is known to be polytopal.

\begin{theorem}[{\cite[Thm.~26]{HohlwegPilaudStella}}]
\label{thm:gvectorFan}
For any finite type exchange matrix~$\B_\circ$, the $\b{g}$-vector fan~$\gvectorFan[\B_\circ]$ is the normal fan of the polytope defined by...
\vincent{TODO}
\end{theorem}

@@

\begin{remark}
\label{rem:exchangeablePairsNKC}
In view of \cref{prop:exchangeablePairsNKC}, it is tempting to look for a characterization of the exchangeable pairs~$\omega, \omega'$ using the kisses between~$\omega$ and~$\omega'$.
However, as illustrated in \cref{fig:nonkissingFans}\,(right), note that
\begin{itemize}
\item two exchangeable walks may kiss along more than one string, but only one is the distinguished string,
\item two non-exchangeable walks can kiss along more than one distinguishable string,
\item two walks that kiss along a single distinguishable string are not always exchangeable,
\item not all strings are distinguishable.
\end{itemize}
\vincent{More explicit. TODO}
\end{remark}

@@

@@

In~\cite{Auslander1984}, M.~Auslander showed that the relations in the Grothendieck group of an Artin algebra are generated by the ones given by almost-split sequences precisely when the algebra is of finite representation type.  Inspired by this result, we will show that a similar statement holds (Theorem~\ref{theo::relations-g-vecteurs}) for the~$\b g$-vectors (or indices) of objects in a~$2$-Calabi--Yau triangulated category.

@@

%%
\subsubsection{$2$-term complexes of projective modules}\label{sect::Kbproj}

@@

%%
\subsubsection{Associative algebras}
